\begin{center}
    \doublespace
    ACKNOWLEDGEMENTS
\end{center}

I am deeply grateful for the guidance I received from my research advisor, Professor Everett Shock. He introduced me to the wonderful and mysterious biological and geochemical diversity of Yellowstone and challenged me to be the creative driving force of my own research. His enthusiasm, encouragement, and scientific insight shaped my transition from student to scientist. I am proud to be a member of his research group.

My advisory committee members Professors Hilairy Hartnett and Pierre Herckes have my deepest gratitude for their thoughtful feedback, hard-hitting questions, and attention to detail.

I extend many thanks to Professor Roger Summons for graciously allowing me to analyze lipids with the instrumentation in his laboratory at the Massachusetts Institute of Technology. I am indebted to Emily Matys for her assistance while running calibration curves and to Dr. Xiaolei Liu for kindly re-running a batch of lipid samples for me after I returned to Arizona. Dr. Florence Schubotz, now a researcher at the University of Bremen, was instrumental in providing analytical instruction, preparing authentic standards, and offering me helpful advice. Her work on lipid biogeochemistry continues to be a source of great inspiration. You have my thanks.

Dr. Alysia Cox and Dr. Jeff Dick, former postdocs of the Group Exploring Organic Processes In Geochemistry (GEOPIG), have my thanks as well. Alysia provided academic guidance that helped shape the organization of content in this dissertation. Jeff introduced me to coding in the R programming language and the R package he developed, CHNOSZ. His enthusiasm for integrating features into CHNOSZ and providing detailed answers to theoretical and technical questions proved extremely helpful.

There are a great many folks who contributed to the collection, processing, and analysis of the Yellowstone samples described in this dissertation. Most of these contributors are fellow members, alumni, or collaborators of GEOPIG at Arizona State University. First and foremost, I thank Jade Woods for her tireless assistance with lipid extractions over the span of six semesters. The work ethic, dependability, and professionalism that she displayed as a volunteer undergraduate research assistant made it a pleasure to work with her. Jenna Donatelli performed temperature, pH, and conductivity measurements, kept cartography records of the hot springs we visited, and lifted spirits with her upbeat attitude. Apar Prasad, Carlos Castellanos, and Kristin Johnson performed field spectrophotometry for water samples at Yellowstone National Park. As one of the few members of GEOPIG with an interest in thermodynamic calculations involving biomolecules, Apar has been a kindred spirit, and as such, his feedback has been valuable. Carlos is a great friend, and I'll never forget jamming out with him to music in the van on our 2011 and 2012 Yellowstone trips. Kristin and I joined GEOPIG at the same time, shared an interview with Everett Shock during our graduate school visitation, and ever since, Kristin has been a close friend. I have fond memories of exploring Tokyo, camping in Greer, Arizona, and trying out new board games with Kristin. Kris Fecteau managed, coordinated, and troubleshot ion chromatography, or in other words, did most of the heavy-lifting for the analysis of major dissolved ions. He also stomped me at chess in Yellowstone. Randall ``Vince" Debes and Kirt Robinson had a great many responsibilities in the field and in the laboratory, including (but not limited to), collecting water and dissolved gas and preparing samples for analysis of major ions and dissolved organic/inorganic carbon (DIC/DOC). In addition to the countless roles they had in the successful analysis of water samples, Vince and Kirt have been inspirational friends and colleagues. I value the years of adventures we have had inside and outside of work. Alta Howells, Michelle Santana, and Brian St. Clair deserve many thanks for analyzing dissolved gas and preparing samples for ion chromatography. I thank Tucker Ely and James Leong for their help with aqueous geochemical speciation; between the three of us, we created an automated tool with a custom thermodynamic database that greatly reduced the time it took to perform several crucial calculations. I also thank Panjai Prapaipong for leading the charge on trace metal analyses, and Natalya Zolotova for her guidance on DIC/DOC analyses. I am grateful for having the chance to work with you all.

I am deeply appreciative of my parents, Clayton and Lisa Boyer, for the love and encouragement they have shown me throughout my life. I consider myself fortunate to have been raised by parents from scientific and medical backgrounds. The knowledge they shared fostered my curiosity about science and the natural world. I also thank my parents-in-law, ``Banana Joe" Joe and Cynthia Halasey, for their love and cheer through the years.

At the time of this writing, Angela Halasey and I have been together for fifteen years and married for six. Angela's unwavering love and support have shaped me in countless positive ways. I am incredibly fortunate to share my life with someone so wonderful.
