\chapter{Thermodynamic equations of state}\label{app4}

\clearpage

\singlespace

% revised HKF equation
\begin{align} \label{eq:HKF_equation}
\begin{split}
\Delta G^{\circ}_{aq} = & \Delta_{f}G^{\circ}_{aq} - S_{aq}^{\circ}(T - T_{r}) - c_{1}\Big[T \cdot \ln{\frac{T}{T_{r}}} - T + T_{r} \Big] \\
      - & c_{2}\Big\{\Big[\Big(\frac{1}{T - \theta}\Big) - \Big(\frac{1}{T_{r} - \theta}\Big) \Big] \Big( \frac{\theta - T}{\theta} \Big) - \frac{T}{\theta^{2}} \ln{\Big( \frac{T_{r}(T - \theta)}{T(T_{r} - \theta)} \Big)} \Big\} \\
      + & a_{1}(P - P_{r}) + a_{2}\ln{\Big(\frac{\Psi + P}{\Psi + P_{r}} \Big)} + \Big( \frac{1}{T - \theta} \Big)\Big[ a_{3}(P - P_{r}) + a_{4}\ln{\Big( \frac{\Psi + P}{\Psi + P_{r}} \Big)} \Big] \\
      + & \omega \Big( \frac{1}{\epsilon} - 1 \Big)
\end{split}
\end{align}
\captionedequationset{The revised Helgeson-Kirkham-Flowers (HKF) equation of state used in this work was adapted from Equation B25 in \cite{shock1992calculation}. Solvation terms not expected to have a significant effect on results within the range of pressure and temperature considered in this study (1 bar and $<$ 100$^{\circ}$C) were omitted. $\Delta G^{\circ}_{aq}$ indicates the apparent standard partial molal Gibbs free energy of formation of the aqueous species from the elements at the temperature and pressure of interest. $S_{aq}^{\circ}$ and $\Delta_{f}G^{\circ}_{aq}$ stand for the standard partial molal entropy of the species and Gibbs free energy of formation of the aqueous species from the elements at 298.15K and 1 bar, respectively. $a_{1}$, $a_{2}$, $a_{3}$, $a_{4}$, $c_{1}$, $c_{2}$, and $\omega$ designate HKF equation of state parameters that are species-dependent. $T$ and $P$ stand for the temperature and pressure of interest. $T_{r}$ and $P_{r}$ represent the reference temperature and pressure of 298.15 K and 1 bar. $\theta$ and $\Psi$ designate solvent parameters equal to 228 K and 2600 bar, respectively. $\epsilon$ stands for the dielectric constant of water at the temperature and pressure of interest.}

\doublespace
\clearpage