\chapter[CONCLUSION]{Conclusion}

\section{Summary}

Organisms require functioning lipids across the entire range of conditions where life is found. As a result, there is an enormous variety of lipid adaptations in biology. As was discussed in this dissertation, however, certain lipid structures are known to function effectively in very different sets of environmental conditions; \textit{e.g.}, glycerol dialkyl glycerol tetraethers (GDGTs) can create stable lipid membranes in both thermophiles and mesophiles \citep{schouten2000widespread}. Likewise, very different sets of lipid distributions may function effectively in the similar environmental conditions; \textit{e.g.}, GDGTs or non-GDGTs can create stable lipid membranes in mesophiles. Any given lipid adaptation, whether it is a double bond, a longer alkyl chain, or a particular headgroup, will have an energetic cost associated with its biosynthesis. These costs are predicted to change depending on the temperature, pressure, and chemical context of the surroundings.

This work provided evidence to support the hypothesis that natural selection favors convergence upon lipid compositions that maximize function while also minimizing energetic cost. First, the abundance-weighted Z\textsubscript{C} of lipids sampled from microorganisms from four alkaline hot springs was shown to correlate positively with the oxidation state of the environment. Adaptations to provide alkyl chain functionality and membrane fluidity were themselves oxidized under oxidized conditions and reduced under reduced conditions, which I hypothesized was energetically cost-efficient. A subsequent thermodynamic analysis provided evidence to support this hypothesis for intact polar lipid (IPL) alkyl chains by predicting that observed IPL distributions were stable relative to each other along redox gradients. Additionally, values of Eh predicted to result in favorable alkyl chain distributions were predicted to be correlated positively with environmental Eh. Lipid headgroups were also investigated and yielded results similar to those of alkyl chains; that hot spring Eh was correlated positively with Eh predicted by energetically stable configurations of observed headgroup ratios. In sum, this work suggests structural adaptations in lipid distributions are not only present because they function for their source organisms, but because they are cost-effective to produce within the temperature and geochemical context of their surroundings.

The thermodynamic studies described in this dissertation did not include calculations involving individual reactions of biosynthesis pathways or intracellular concentrations of chemical species in their evaluation of lipid cost. Instead, net reactions to form lipids from measured concentrations of bioavailable solutes alone provided evidence that observed thermophile lipid distributions may be energetically cost-effective. It is essential for cells to regulate intracellular conditions such as pH and solute concentrations for metabolic pathways to proceed. Maintaining chemical disequilibrium from the external environment requires energy. However, I would argue this energy requirement is based on how well a cell is adapted to its surroundings; energetic shortcuts to synthesize necessary biomolecules from available materials in the surroundings would offset a portion of this energy requirement. This raises the possibility that conditions \textit{outside} of the cell matter more for the energetic costs of biomolecules than the conditions \textit{inside} the cell. This concept could serve as a foundation for studies in molecular biology to evaluate potential costs of biomolecules synthesized by microbial communities in a variety of natural systems and in culture.

\section{Future Work}

Studies of lipid distributions in natural systems could be mainly exploratory (\textit{i.e.}, which lipids are found under which sets of geochemical conditions) but could be supplemented by calculation of Z\textsubscript{C} or thermodynamic predictions to allow quantitative assessment of lipid distributions in terms of energetic cost. These studies could also consider the phylogeny of source organisms, which was not done here. For instance, would a cyanobacteria-dominated microbial community that is adapted to thrive in water rich in hydrogen and methane, such as conditions found in fluids discharged from serpentinizing systems \citep{mccollom2009thermodynamic}, tend to produce lipid compositions with reduced carbon relative to cyanobacteria-dominated communities in oxidized ocean surface water? How phylogenetically related can two microbial communities be while having drastically different lipid compositions, and how might functional necessity and energetic cost help explain these differences?

These types of questions could also be probed in laboratory experiments. Various growth media of known chemical compositions could be inoculated with a mixed culture to isolate sub-sets of microbes that grow best under those conditions. Thermodynamic calculations based on those described in this work might then be used to predict whether the lipid compositions of isolated sub-sets tend to be energetically favorable in the chemical conditions of their own growth media relative to all others in the experiment. Would thermodynamic predictions offer greater insight into the expression of observed lipid distributions in these various treatments than the phylogenetic compositions of the isolated microbial sub-sets?

Future studies should attempt to be as quantitative as possible when measuring lipid abundances to allow estimation of weighted molecular properties like the average oxidation state of carbon (Z\textsubscript{C}) or partial molal thermodynamic properties such as Gibbs free energies of formation, volumes, entropies, and the like. The power to describe lipid distributions with of Z\textsubscript{C} or thermodynamic calculations depends on the physical and chemical parameters of the system. As such, I recommend that these studies strive to be as complete as possible when measuring temperature, pressure, and chemical composition of the natural system or growth media.

Thermodynamic calculations could explore the formation of lipids from a different set of basis species than the `autotrophic' set used in this work. It may be more appropriate to use acetate, for instance, to predict lipid distributions in a predominantly heterotrophic microbial community. Future studies could focus on reducing sources of error to aid in the comparison of thermodynamically predicted and analytically observed lipid abundances. Constraining sources of error introduced by lack of authentic response factors may offer a better estimation of aminopolyol BHP abundance. Alternate estimation schemes could be explored to estimate lipid thermodynamic properties. Liquid-phase or crystalline-phase alkyl chains, or a combination of the two, could be included in formation reactions from aqueous basis species to simulate the formation of a membrane lipid; this may offer a better predictive model than the formation of free aqueous lipids. Alternative methods for predicting complex distributions of lipids from basis species could be explored, and results could be compared to outcomes produced by the hypothetical site-averaged chains.

% extras, if needed:
% Lipid studies addressing "maximization of function". Assumed, but not tested here. Some studies have addressed function (e.g. SQ headgroups and ability to photosynthesize, or knockout membranes with no BHPs, etc.)

\section{Broader Impacts}

Interpreting the expression of lipids in terms of functional benefit and energy cost has the potential to change our understanding of the evolution of lipids, geochemical controls on lipid distributions, and ultimately how biomarkers are interpreted in the rock record.

Biochemistry emerged from geochemical processes, though it is still unclear when lipids emerged in the history of life. Some theories posit that lipids were essential for metabolic encapsulation, growth, and division in the first proto-organisms \citep{deamer2002first, luisi2016emergence}. Placing abiotic lipid synthesis in a thermodynamic framework could help narrow down geochemical conditions conducive to spontaneous lipid formation that could then be reconciled with the conditions proposed to allow early growth and division. It has also been theorized that lipid membranes evolved later.  These theories hold that the last universal common ancestor (LUCA) that led to modern Bacteria, Archaea, and Eukarya lacked lipid membranes \citep{koga1998did, martin2003origins}, or had membranes with simpler lipid compositions \citep{sojo2014bioenergetic}. The notion that biology maximizes lipid function and minimizes energy cost could be applied to aid understanding of evolutionary pressures imposed by geochemistry that led to the divergence of lipid synthesis pathways among prokaryotes and eukaryotes after LUCA. While these are meant to represent specific examples where lipids are the focus, I am optimistic that thermodynamic assessments of the energetic costs of biomolecules would benefit evolutionary studies in general.

% geochemical controls on lipid distributions. Frame in astrobiological sense. Function of new lipid structures still under invest, but thermo predictions could help uncover sources.

Lipid biomarkers have great potential to reveal detail about the environmental conditions experienced by their source communities, as well as geochemistry experienced by sedimentary rocks after diagenesis. In \cite{ourisson1979hopanoids}, the authors report complex distributions of hopanoid lipid degradation products from a variety of sediments. They reported that hundreds of chemical structures were observed in these biomarkers, ranging from carboxylic acids to aldehydes to hydrocarbons to aromatic compounds, and emphasized that a wealth of information about geologic history could be unlocked if more could be learned about geochemical influences on the transformation of biomarkers. A review by \cite{ourisson1992hopanoids} included speculation about patterns in hopanoid degradation products expected to form under oxidized and reduced conditions, but without a thermodynamic framework, these patterns can only be interpreted qualitatively. This dissertation focused on thermal and geochemical controls on lipid expression in active, living systems, but thermodynamic data could be expanded to include biomarker degradation products. Thermodynamic calculations could potentially revolutionize the interpretation of biomarkers by allowing quantitative predictions regarding the chain of geologic events that led to observed biomarker distributions in the rock record.

\section{Concluding Remarks}
Geochemistry dictates what is necessary for habitability. This dissertation explored what that means for lipid adaptation; a lipid composition must function within the conditions experienced by the microbial community that produced it, and it must also have an acceptable energetic cost. This energetic cost is dictated by the temperature, pressure, and chemical composition of the system. A microbial community adapted to its surroundings is one that has evolved lipid distributions that are both functional and cost-effective. The composition of lipid structures produced by a microbial community should therefore reflect this interplay of energetic cost and functional benefit. Energetic cost can be assessed quantitatively, but this requires a thermodynamic framework to work within. This dissertation provides a step in that direction. Future work aimed at developing this framework further has the potential to shed great insight into why life produces the chemical structures that it does; not only with respect to lipid adaptation in thermophiles but for biomolecular adaptation across all of biology.

% how lipid composition in eukaryotic organelles may have evolved over time.

% paleothermometry - paleoredox

%Everett wrote for NSF proposal: The development of thermodynamic data for lipid biomolecules and their biomarker degradation products will unlock the record of geochemical transformations recorded in these compounds, which will test, confirm, and challenge conclusions drawn from mineralogical, isotopic, and trace element datasets from sedimentary rocks. It is possible that biomarkers will become as useful for deciphering the geological history of sedimentary rocks as they are presently for determining the depositional environments of the original sediments.
%Biomarkers hold the potential to reveal extraordinary detail about past environmental conditions that prevailed anywhere that sediments accumulate and preserve geobiochemical information. There is an impressive potential for a comprehensive understanding of global change through integrating biomarkers with many other lines of evidence from the geological record. Remarkably, biomarker research is predominantly an analytical pursuit with few experimental studies and even fewer efforts at a theoretical framework. This project will provide a foundation for that framework, which will yield novel applications in microbial physiology, evolution of biochemical pathways, climate change, and the emerging field of geobiochemistry


% Concluding remarks suggested by committee:
% Potential applications: interpretation of lipid distributions in the rock record and their controls and eventually interpreting how lipid distributions are altered after deposition; Predicting which lipid biomarkers to target in astrobiology missions; New insight into the evolution of lipids in biology. Future work could compare lipid thermo calcs to (meta)genomes to show that bioenergetic trends exist regardless of organisms.

% astrobiology impact: cite paper about lipid patterns in astrobio I sent to Melissa.

